### 2025-04-04 GettingStarted.Rmd
install and test with vignettes

---
title: "WHAM MSE startup"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Following the vignettes from Chengxue Li at https://lichengxue.github.io/whamMSE/

Installed as instructed under https://lichengxue.github.io/whamMSE/01.Installation.html but WHAM dev branch already installed

Running self test and cross test https://lichengxue.github.io/whamMSE/02.Self-test-and-Cross-test.html

Simple self test need to change some inputs to avoid error, code modified from vignette below, all other code blocks from vignette

```{r}
year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 0     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 2,
                            n_regions = 2,
                            n_indices = 2,
                            n_fleets = 2,
                            n_seasons = 4,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "medium",
                            n_ages = 12,
                            #Fbar_ages = 12,
                            #recruit_model = 2,
                            F_info = list(F.year1 = 0.2, Fhist = "constant", Fmax = 0.4, Fmin = 0.2, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625,
                            fracyr_seasons = NULL,
                            fleet_regions = NULL, # was fleet_pointer
                            index_regions = NULL, # was index_pointer
                            user_waa = NULL,
                            user_maturity = NULL,
                            bias.correct.process = FALSE,
                            bias.correct.observation = FALSE,
                            bias.correct.BRPs = FALSE,
                            mig_type = 0,
                            XSPR_R_opt = 2,
                            move_dyn = 0, # natal homing is assumed
                            onto_move = 0,
                            onto_move_pars = NULL,
                            apply_re_trend = 0,
                            trend_re_rate = NULL,
                            apply_mu_trend = 0,
                            trend_mu_rate = NULL, 
                            age_mu_devs = NULL)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info
```

Which option to run? Trying 3a bi-directional

```{r}
basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 2)
# Note: the default is "bidirectional" movement (e.g. stock 1 move to region 2 and stock 2 move to region 1)

move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3, 0.1), move.re = "constant") # Note: default is move = 0.3 (constant) for stock1 and 0.1 (constant) for the other stocks
```

Selectivity and M, 0.2 of course

```{r}
n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
```

NAA

```{r}
sigma        <- "rec+1"
re_cor       <- "iid"
ini.opt      <- "equilibrium" # option <- c("age-specific-fe", "equilibrium")
Rec_sig      <- 0.5 # (sigma for recruitment)
NAA_sig      <- 0.5 # (sigma for NAA)

sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages
sigma_vals[,,1] = Rec_sig

# Set initial NAA for each stock
log_N1    <- rep(10,n_stocks)
log_N1[1] <- log(exp(10)*2) # N1_stock1 is 2 times higher than N1_stock2
N1_pars   <- generate_ini_N1(log_N1,basic_info,ini.opt)

# Set mean recruitment para. for each stock
mean_rec_par <- list()
for (i in 1:n_stocks) mean_rec_par[[i]] <- exp(log_N1[i])

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 2,  # rec random around the mean
               recruit_pars = mean_rec_par, 
               sigma_vals = sigma_vals,  
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where) # NAA_where must be specified in basic_info!
```

WHAM input

```{r}
input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)
```

OM generation

```{r}
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)
```

Self test

```{r}
sim_fn <- function(om, self.fit = FALSE){
  input <- om$input
  input$data <- om$simulate(complete=TRUE)
  # $simulate() is a function to generate datasets given the parameters defined in the OM
  # Note: process error and observation error are both turned on when simulating datasets 
  # To turn them off, use basic_info$simulate_process_error = 0 and basic_info$simulate_observation_error = 0
  if(self.fit) {
    fit <- fit_wham(input, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = FALSE)
    return(fit)
  } else return(input) 
}
# Note: turn on do.osa to calculate one-step-ahead residuals
# Note: turn on do.retro to calculate retrospective bias
set.seed(12345)
self_sim_fit <- sim_fn(om, self.fit = TRUE)
```

Check convergence (pass)

```{r}
check_convergence(self_sim_fit) # check the model convergence
```

Cross test: setup

```{r}
# EM with different NAA configuration
sigma      <- "rec+1"
re_cor     <- "2dar1" # Misspecification here

ini.opt    <- "equilibrium"
NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               NAA_where = basic_info$NAA_where)

# Generate wham input
input <- prepare_wham_input(basic_info = basic_info, selectivity = sel, M = M, NAA_re = NAA_re, move = move,
                            catch_info = catch_info, index_info = index_info, F = F_info)

# Generate EM
em <- fit_wham(input, do.fit = F, do.brps = F, MakeADFun.silent = TRUE)
```

Cross test: perform

```{r}
sim_fn2 <- function(om, em, cross.fit = FALSE){
  input <- em$input
  input$data <- om$simulate(complete=TRUE)
  if(cross.fit) {
    fit <- fit_wham(input, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = FALSE)
    return(fit)
  } else return(input) 
}

set.seed(12345)
cross_sim_fit <- sim_fn2(om, em, cross.fit = TRUE)
```

Cross test: convergence test (fails)

```{r}
check_convergence(cross_sim_fit)
```
Replicates

Self test

```{r}
nsim <- 100 
set.seed(8675309) 
sim_input <- list()
sim_input <- lapply(1:nsim, function(x) {
  input_i <- om$input
  sim <- om$simulate(complete=TRUE)
  input_i$data <- sim
  return(input_i)
})

# Self test
sim_fits <- list()
sim_fits <- lapply(1:nsim, function(x){
  cat(paste("model_fit:", x, "start \n"))
  out <- fit_wham(sim_input[[x]], do.osa = FALSE, MakeADFun.silent = TRUE, retro.silent = TRUE, save.sdrep = FALSE)
  cat(paste("model_fit:", x, "done \n"))
  return(out)
})

# Summarize results 
conv <- sapply(1:nsim, function(x){
  if (length(sim_fits[[x]]) != 0) {
    if (sim_fits[[x]]$is_sdrep & !sim_fits[[x]]$na_sdrep & !sim_fits[[x]]$hessian) {
      conv = TRUE } else conv = FALSE
  } else conv = FALSE
  return(conv)
})
cat(paste("Convergence rate:", sum(conv)/nsim))

mean_rec_par <- lapply(1:nsim, function(x){
  mean_rec_par_est <- sim_fits[[x]]$parList$mean_rec_pars[,1]
  mean_rec_par_true <- sim_fits[[x]]$input$par$mean_rec_pars[,1]
  mean_rec_par <- cbind(mean_rec_par_est,mean_rec_par_true)
  return(mean_rec_par)
})
print(mean_rec_par)

SSB <- lapply(1:nsim, function(x){
  SSB_est <- sim_fits[[x]]$rep$SSB
  SSB_true <- sim_fits[[x]]$input$data$SSB
  SSB <- cbind(SSB_est, SSB_true)
  return(SSB)
})
print(SSB)
```

### 2025-04-07 GettingStarted.Rmd
start pages

@@
 Installed as instructed under https://lichengxue.github.io/whamMSE/01.Installation.html but WHAM dev branch already installed
+
+```{r}
+library(wham)
+library(whamMSE)
+
+main.dir = here::here()
+```
 
@@
 
-Self test
+Self test, run once, took 6 hours
 
-```{r}
+```{r, eval=FALSE}
 nsim <- 100 
@@
 })
+
+#skg added
+saveRDS(sim_fits, here::here("localhugefiles/sim_fits_self_test.rds"))
 
@@
 print(SSB)
+
 ```
+
+What does the object look like? Huge
+
+```{r}
+sim_fits <- readRDS(here::here("localhugefiles/sim_fits_self_test.rds"))
+
+```

### 2025-04-07 docs/index.Rmd
start pages

---
title: "WHAM MSE Pages"
author: "Sarah Gaichas"
output: html_document
---

<button>
    <a href="https://github.com/NOAA-EDAB/WHAMmse_longlived"> WHAM mse longlived github repo </a>
</button>

<!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

## Analysis pages

<a href="https://noaa-edab.github.io/WHAMmse_longlived/GettingStarted.html" target="_blank"> Test WHAM MSE code from vignettes - April 2025 </a>

<a href="https://noaa-edab.github.io/WHAMmse_longlived/WorkflowDecisions.html" target="_blank"> Documenting long lived MSE workflow - April 2025 </a>

<a href="https://noaa-edab.github.io/WHAMmse_longlived/Results.html" target="_blank"> Initial results - April 2025 </a>

### 2025-04-07 MSEsetup.Rmd
trying long lived MSE setup, currently breaking

---
title: "MSE Setup"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output:
  html_document:
    code_fold: hide
link-citations: yes
csl: "canadian-journal-of-fisheries-and-aquatic-sciences.csl"
bibliography: MSE.bib
urlcolor: blue  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

Following the MSE vignette from Chengxue Li at https://lichengxue.github.io/whamMSE/

```{r}
library(wham)
library(whamMSE)

main.dir = here::here()
```

## Run the vignette exactly first, breaks

```{r, eval=FALSE}
year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 3     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 2,
                            n_regions = 2,
                            n_indices = 2,
                            n_fleets = 2,
                            n_seasons = 4,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "medium",
                            n_ages = 12,
                            #Fbar_ages = 12,
                            #recruit_model = 2,
                            F_info = list(F.year1 = 0.2, Fhist = "F-H-L", Fmax = 2, Fmin = 1, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info

basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 2) # "bidirectional" movement

move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3,0.1), 
                      move.re = "iid_y", move.sigma = 0.5, prior.sigma = 0.5, use.prior = TRUE)

n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))

sigma      <- "rec+1"
re_cor     <- "iid"
ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")

# Set para. for B-H function
alpha <- 12
beta  <- 1.5e-4

# Set sigma for NAA
NAA_sig <- 0.2
sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"

# Set initial NAA for each stock
log_N1  <- rep(10, n_stocks) # Create difference between stocks
N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 3,
               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
               sigma_vals = sigma_vals,
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where)

# recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
# recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
# recruit_model = 3: Beverton-Holt stock-recruitment with annual recruitments as random effects
# recruit_model = 4: Ricker stock-recruitment with annual recruitments as random effects

# 1. recruit_pars: a list (length = n_stocks) of vectors of initial parameters for recruitment model. 
# If $recruit_model is 3 (B-H) or 4 (Ricker), parameters are "alpha" and "beta".

# 2. sigma_vals: Initial standard deviation values to use for the NAA deviations. Values are not used if recruit_model = 1 
# If sigma="rec": must be a list (length = n_stocks) of single values
# If sigma="rec+1": a list (length = n_stocks) of 2 values must be specified. First is for the first age class (recruits), second is for all other ages.

input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)

random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)

om_with_data <- update_om_fn(om, seed = 123, random = random)

assess.interval <- 3 # Note: assessment interval is 3 years, given the feedback period is 3 years, there will be only 1 assessment
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs

n_stocks = n_regions = n_fleets = n_indices = 1

sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = TRUE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            aggregate_index_info = list(index_pointer = rep(1,2),
                                                        use_catch_weighted_waa = TRUE),
                            aggregate_catch_info = list(fleet_pointer = rep(1,2),
                                                        use_catch_weighted_waa = TRUE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            hcr = list(hcr.type = 1, hcr.opts = NULL), 
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE)
```

## Modify for long-lived, this runs but crashes

Basic info for a long-lived species. Based on @wiedenmann_evaluation_2017, Table 3 B and C.

A long-lived species recruits at 5 years, lives to 20, mean M is 0.1, matures at 7.

Modify from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html

Start with a 20 year burn in and 20 MSE years. First try constant low F. Then work up to complex case.

Fishing pattern is "updown" increasing to Fmax then decreasing back to Fmin

```{r}

year_start  <- 1  # starting year in the burn-in period
year_end    <- 20  # end year in the burn-in period
MSE_years   <- 20     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation

info <- generate_basic_info(n_stocks = 1, #2,
                            n_regions = 1, #2,
                            n_indices = 1, #2,
                            n_fleets = 1, #2,
                            n_seasons = 4,
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "long",
                            n_ages = 20,
                            #Fbar_ages = 12,
                            #recruit_model = 2,
                            F_info = list(F.year1 = 0.01, Fhist = "constant", Fmax = 2, Fmin = 0.5, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = 0.625, q = 0.2),
                            fracyr_spawn = 0.625)

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info
```

Movement, do we need this? 1 stock 1 region model, turn it off

```{r}
basic_info <- generate_NAA_where(basic_info = basic_info, move.type = NULL) # "bidirectional" movement

move <- NULL

#move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3), #,0.1
#                     move.re = "iid_y", move.sigma = 0.5, prior.sigma = 0.5, use.prior = TRUE)
```
Adjust for 1 area 1 stock, long lived M=0.1

```{r}
n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages)))
```

```{r}
sigma      <- "rec+1"
re_cor     <- "iid"
ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")

# Set para. for B-H function
alpha <- 12
beta  <- 1.5e-4

# Set sigma for NAA
NAA_sig <- 0.2
sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"

# Set initial NAA for each stock
log_N1  <- rep(10, n_stocks) # Create difference between stocks
N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 3,
               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
               sigma_vals = sigma_vals,
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where)

# recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
# recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
# recruit_model = 3: Beverton-Holt stock-recruitment with annual recruitments as random effects
# recruit_model = 4: Ricker stock-recruitment with annual recruitments as random effects

# 1. recruit_pars: a list (length = n_stocks) of vectors of initial parameters for recruitment model. 
# If $recruit_model is 3 (B-H) or 4 (Ricker), parameters are "alpha" and "beta".

# 2. sigma_vals: Initial standard deviation values to use for the NAA deviations. Values are not used if recruit_model = 1 
# If sigma="rec": must be a list (length = n_stocks) of single values
# If sigma="rec+1": a list (length = n_stocks) of 2 values must be specified. First is for the first age class (recruits), second is for all other ages.
```

Make inputs for the OM

```{r}
 input <- prepare_wham_input(basic_info = basic_info, 
                            selectivity = sel, 
                            M = M, 
                            NAA_re = NAA_re, 
                            move = move,
                            catch_info = catch_info, 
                            index_info = index_info, 
                            F = F_info)
```

Make the OM

```{r}
random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)
```

One dataset

```{r}
om_with_data <- update_om_fn(om, seed = 123, random = random)
```

Assessment specifications, every 3 years as a start, MP could be annual

```{r}
assess.interval <- 3 # 
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs
```

Try a simple EM, looks like EM1 [here](https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#10_Description_of_assessment_models:)

```{r}
# this matches the OM
n_stocks = n_regions = n_fleets = n_indices = 1

sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = TRUE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            aggregate_index_info = list(index_pointer = rep(1,1),
                                                        use_catch_weighted_waa = TRUE),
                            aggregate_catch_info = list(fleet_pointer = rep(1,1),
                                                        use_catch_weighted_waa = TRUE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            hcr = list(hcr.type = 1, hcr.opts = NULL), 
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE)
```

# References

### 2025-04-08 MSEsetup.Rmd
loop_through_fn not working for either case

@@
                             #recruit_model = 2,
-                            F_info = list(F.year1 = 0.01, Fhist = "constant", Fmax = 2, Fmin = 0.5, change_time = 0.5, user_F = NULL),
+                            F_info = list(F.year1 = 0.01, Fhist = "constant", Fmax = 0.01, Fmin = 0.01, change_time = 0.5, user_F = NULL),
                             catch_info = list(catch_cv = 0.1, catch_Neff = 100),

### 2025-04-11 MSEsetup.Rmd
Cheng fixed MSE test, now runs

--- 2025-04-07 MSEsetup.Rmd
@@
 
-## Run the vignette exactly first, breaks
+## Run the vignette exactly first, Cheng fixed! works now
 
@@
                sigma_vals = sigma_vals,
-               N1_pars = N1_pars,
-               NAA_where = basic_info$NAA_where)
+               N1_pars = N1_pars)#,
+               #NAA_where = basic_info$NAA_where)
 
@@
                                           do.move = FALSE, est.move = FALSE),
-                            aggregate_index_info = list(index_pointer = rep(1,2),
-                                                        use_catch_weighted_waa = TRUE),
-                            aggregate_catch_info = list(fleet_pointer = rep(1,2),
-                                                        use_catch_weighted_waa = TRUE),
+                            # Please use this code:
+                            aggregate_catch_info = list(n_fleets = 1, 
+                                                        fleet_pointer = c(1,1), 
+                                                        use_catch_weighted_waa=TRUE, 
+                                                        catch_Neff = 100, 
+                                                        catch_cv = 0.1),
+                            # Please use this code:
+                            aggregate_index_info = list(n_indices = 1, 
+                                                        index_pointer = c(1,1), 
+                                                        use_catch_weighted_waa=TRUE, 
+                                                        index_Neff = 100, 
+                                                        index_cv = 0.1),
                             assess_years = assess.years, 
@@
                             save.last.em = TRUE)
+
+mods[[1]]$om$rep$pred_catch # Check catch in the feedback loop
+mods[[1]]$om$rep$SSB  # Check SSB in the feedback loop
 ```
 
-## Modify for long-lived, this runs but crashes
+## Modify for long-lived, Cheng fixed and now runs
 
@@
 
-Fishing pattern is "updown" increasing to Fmax then decreasing back to Fmin
+Fishing pattern is "constant" at first to get running
 
@@
                             n_fleets = 1, #2,
-                            n_seasons = 4,
+                            n_seasons = 1, # don't need seasons
                             base.years = year_start:year_end,
@@
                             n_ages = 20,
-                            #Fbar_ages = 12,
-                            #recruit_model = 2,
-                            F_info = list(F.year1 = 0.01, Fhist = "constant", Fmax = 2, Fmin = 0.5, change_time = 0.5, user_F = NULL),
+                            # We want a bit higher F in the historical period
+                            F_info = list(F.year1 = 0.1, Fhist = "constant", Fmax = 0.1, Fmin = 0.1, change_time = 0.5, user_F = NULL),
                             catch_info = list(catch_cv = 0.1, catch_Neff = 100),
@@
                             F = F_info)
+
+# IMPORTANT!#
+# This appears to be due to the initial F value for the Newton iterations to find F40. The default value is too high for long-lived fish
+input$data$FXSPR_init[] <- 0.01 # change to a low value
+
 ```
@@
 # this matches the OM
-n_stocks = n_regions = n_fleets = n_indices = 1
-
-sel_em <- list(model=rep("logistic",n_fleets+n_indices),
-               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
-
+# n_stocks = n_regions = n_fleets = n_indices = 1
+# 
+# sel_em <- list(model=rep("logistic",n_fleets+n_indices),
+#                initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
+# 
 NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
-
-M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M
+# 
+# M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M
 
@@
                             random = random,
-                            M_em = M_em, 
-                            sel_em = sel_em, 
-                            NAA_re_em = NAA_re_em, 
+                            M_em = M, # use OM M
+                            sel_em = sel, # use OM sel
+                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                             move_em = NULL,
                             age_comp_em = "multinomial",
-                            em.opt = list(separate.em = TRUE, separate.em.type = 1, 
+                            # Here is the correct code: separate.em = FALSE also works for one-area model
+                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                           do.move = FALSE, est.move = FALSE),
-                            aggregate_index_info = list(index_pointer = rep(1,1),
-                                                        use_catch_weighted_waa = TRUE),
-                            aggregate_catch_info = list(fleet_pointer = rep(1,1),
-                                                        use_catch_weighted_waa = TRUE),
                             assess_years = assess.years, 
@@
                             year.use = 20, # number of years of data you want to use in the assessment model
-                            hcr = list(hcr.type = 1, hcr.opts = NULL), 
                             seed = 123,
                             save.sdrep = FALSE, 
-                            save.last.em = TRUE)
+                            save.last.em = TRUE,
+                            FXSPR_init = 0.01) # IMPORTANT!
 ```

### 2025-04-11 MSEsetup.Rmd
adds degraded future surveys

@@
 
+Now set up as in new Single Stock MSE vignette: https://lichengxue.github.io/whamMSE/11.Single-Stock-MSE.html
+
 ```{r}
@@
 
+Try same model with poorer survey in the future
+
+See https://lichengxue.github.io/whamMSE/15.Reduce-Data-Availability-In-Assessment.html#6_Increase_observation_error_and_reduce_sample_size_for_survey_indices_in_the_feedback_period
+
+```{r}
+agg_index_sigma = input$data$agg_index_sigma
+agg_index_sigma[21:40,] = 1 # Increase CV for both survey indices in the feedback period
+index_Neff = input$data$index_Neff
+index_Neff[21:40,] = 30 # Decrease ESS for both survey indices in the feedback period
+
+input <- update_input_index_info(input, agg_index_sigma, index_Neff) # Update input file
+```
+
+Generate new om and dataset, keep same assessment interval as above
+
+```{r}
+om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
+om_with_data <- update_om_fn(om, seed = 123, random = random)
+```
+
+Run new loop
+
+```{r}
+mods[[2]] = loop_through_fn(om = om_with_data,
+                            em_info = info, 
+                            random = random,
+                            M_em = M, # use OM M
+                            sel_em = sel, # use OM sel
+                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
+                            move_em = NULL,
+                            age_comp_em = "multinomial",
+                            # Here is the correct code: separate.em = FALSE also works for one-area model
+                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
+                                          do.move = FALSE, est.move = FALSE),
+                            assess_years = assess.years, 
+                            assess_interval = assess.interval, 
+                            base_years = base.years,
+                            year.use = 20, # number of years of data you want to use in the assessment model
+                            seed = 123,
+                            save.sdrep = FALSE, 
+                            save.last.em = TRUE,
+                            FXSPR_init = 0.01) # IMPORTANT!
+
+```
+
+Visualize 2 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)
+
+```{r}
+
+par(mfrow = c(1,2))
+SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
+plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB") 
+colors <- c("red","green","purple","orange")
+for (i in length(SSB_s1)) {
+  lines(SSB_s1[[i]], col = colors[i-1],lty = i)
+}
+legend("topleft",legend = paste0("EM ", 1:2), col = c("blue",colors),lty=1:5, cex=0.8)
+
+Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
+plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,30000), main = "OM Catch")
+colors <- c("red","green","purple","orange")
+for (i in length(Catch_s1)) {
+  lines(Catch_s1[[i]], col = colors[i-1],lty = i)
+}
+
+```
+
 # References

### 2025-04-11 MSEsetup.Rmd
with Cheng's fixes and additional no survey run

@@
                             year.use = 20, # number of years of data you want to use in the assessment model
+                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                             seed = 123,
@@
                             year.use = 20, # number of years of data you want to use in the assessment model
+                            add.years = TRUE,
+                            update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
                             seed = 123,
@@
 Visualize 2 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)
+
+This compares a good survey in the future (1) with a poor survey in the future (2)
 
@@
 ```
+Now add the one with no survey, from Cheng's code
+
+```{r}
+# ---------------------------------
+# Actually the above section (CV and Neff) will be useless because we will not have survey data in the feedback period), see code below:
+# ---------------------------------
+remove_agg = TRUE # remove a aggregate index for some years
+remove_agg_pointer = 1 # target on index 1
+remove_agg_years = 21:40 # all feedback years
+remove_paa = TRUE # Also remove age comp for that index 
+remove_paa_pointer = 1 # target on index 1
+remove_paa_years = 21:40 # all feedback years
+
+input <- update_input_index_info(input, agg_index_sigma, index_Neff,
+                                 remove_agg, remove_agg_pointer, remove_agg_years,
+                                 remove_paa, remove_paa_pointer, remove_paa_years) # Update input file
+```
+
+Generate new om and dataset, keep same assessment interval as above
+
+```{r}
+om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
+om_with_data <- update_om_fn(om, seed = 123, random = random)
+```
+
+Run new loop
+
+```{r}
+# this simulation actually represents the data-missing scenario.
+mods[[3]] = loop_through_fn(om = om_with_data,
+                            em_info = info, 
+                            random = random,
+                            M_em = M, # use OM M
+                            sel_em = sel, # use OM sel
+                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
+                            move_em = NULL,
+                            age_comp_em = "multinomial",
+                            # Here is the correct code: separate.em = FALSE also works for one-area model
+                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
+                                          do.move = FALSE, est.move = FALSE),
+                            assess_years = assess.years, 
+                            assess_interval = assess.interval, 
+                            base_years = base.years,
+                            year.use = 20, # number of years of data you want to use in the assessment model
+                            add.years = TRUE,
+                            # Must have the below code! Otherwise the EM will assume a constant CV and Neff (same as the CV and Neff in the first year) for the survey
+                            update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
+                            seed = 123,
+                            save.sdrep = FALSE, 
+                            save.last.em = TRUE,
+                            FXSPR_init = 0.01) # IMPORTANT!
+
+```
+
+Visualize 3 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)
+
+```{r}
+
+par(mfrow = c(1,2))
+SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
+plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(1e4,130000), main = "OM SSB",, lwd = 2) 
+lines(SSB_s1[[2]], col = "red",lty = 1, lwd = 2)
+lines(SSB_s1[[3]], col = "orange",lty = 1, lwd = 2)
+
+legend("topleft",legend = c("EM1_good_survey", "EM2_poor_survey","EM3_missing_survey"), col = c("blue","red","orange"),lty=1, cex=0.8)
+
+Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
+plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,15000), main = "OM Catch", lwd = 2)
+lines(Catch_s1[[2]], col = "red",lty = 1, lwd = 2)
+lines(Catch_s1[[3]], col = "orange",lty = 1, lwd = 2)
+
+```
 

### 2025-04-14 Scenarios.Rmd
scenarios

---
title: "MSE Scenarios for short and long lived"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup for MSEs

Single area

Two surveys spring and fall

Testing alternating spring and fall

Baseline is all data all the time,

CV = empirical value from stockeff across short and long lived species

Eff = 100

Catch CV = 10%

Eff = 100

Project at F40%

Assessment frequency 3 years

IID NAA

30 year spin up 30 year loop (10 assessment cycles)

F scenario for spin up? high followed by lower? Yes

"F-H-L" Fmax = 2.5, Fmin = 1, changepoint = 0.5

sigma squared for NAA and recruitment?

cod assessment or redfish for these

## Degraded data quality scenarios:

Alternate spring and fall, double CV and half Eff (50)

Fishery data, double CV (0.2) and half Eff (50)

Keep multinomial

Single fleet logistic selectivity: shift one age to the right from maturity ogive

Different selectivity by spring and fall survey: mimic some long lived species

## Empirical management procedure

Index based method:  Ismooth averaging most recent 3 years of whatever surveys (rescaled to mean of each using Ismooth package)

Ismooth package is here https://github.com/cmlegault/PlanBsmooth

100 replicates for each, try the parallel processing

* Base scenario 
* Alternating surveys and less age comps with 3 year assessments
* Empirical management procedures (later)

### 2025-04-14 Scenarios.Rmd
start modifying script; not done

@@
 
-"F-H-L" Fmax = 2.5, Fmin = 1, changepoint = 0.5
+Double Fmsy for 15 years and then Fmsy for 15 years
+
+F.year1 = 0.1, Fhist = "F-H-L", Fmax = 2.5, Fmin = 1, changepoint = 0.5
 
@@
 cod assessment or redfish for these
+
+I think this script does this
+
+```{r}
+
+library(wham)
+library(whamMSE)
+
+main.dir = here::here()
+
+# get median DOY for each seasonal survey
+# from a local surdat pull
+# fracyr <- survdat_nobio$survdat |> #file from last May, could update
+#   dplyr::mutate(DOY = lubridate::yday(EST_TOWDATE)) |> 
+#   dplyr::select(SEASON, DOY) |> dplyr::distinct() |> 
+#   dplyr::group_by(SEASON) |> 
+#   dplyr::summarise(medDOY = median(DOY), 
+#                    fracyrmed = medDOY/365)
+
+year_start  <- 1  # starting year in the burn-in period
+year_end    <- 30  # end year in the burn-in period
+MSE_years   <- 30     # number of years in the feedback loop
+# Note: no need to include MSE_years in simulation-estimation
+
+info <- generate_basic_info(n_stocks = 1, 
+                            n_regions = 1, 
+                            n_indices = 2, # spring and fall
+                            n_fleets = 1, 
+                            n_seasons = 1, # don't need seasons
+                            base.years = year_start:year_end,
+                            n_feedback_years = MSE_years,
+                            life_history = "long",
+                            n_ages = 20,
+                            # We want a bit higher F in the historical period
+                            F_info = list(F.year1 = 0.1, Fhist = "F-H-L", Fmax = 2.5, Fmin = 1, change_time = 0.5, user_F = NULL),
+                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
+                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = c(0.295, 0.812), q = 0.2),
+                            fracyr_spawn = 0.625)
+
+basic_info = info$basic_info # collect basic information
+catch_info = info$catch_info # collect fleet catch information
+index_info = info$index_info # collect survey information
+F_info = info$F # collect fishing information
+
+basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 3) # no movement
+
+move <- NULL
+
+n_stocks  <- as.integer(basic_info['n_stocks'])
+n_regions <- as.integer(basic_info['n_regions'])
+n_fleets  <- as.integer(basic_info['n_fleets'])
+n_indices <- as.integer(basic_info['n_indices'])
+n_ages    <- as.integer(basic_info['n_ages'])
+
+# Selectivity Configuration
+fleet_pars <- c(5,1)
+index_pars <- c(2,1)
+sel <- list(model=rep("logistic",n_fleets+n_indices),
+            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
+
+# M Configuration
+M <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages)))
+
+sigma      <- "rec+1"
+re_cor     <- "iid"
+ini.opt    <- "equilibrium" # option   <- c("age-specific-fe", "equilibrium")
+
+# Set para. for B-H function
+alpha <- 12
+beta  <- 1.5e-4
+
+# Set sigma for NAA
+NAA_sig <- 0.2
+sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"
+
+# Set initial NAA for each stock
+log_N1  <- rep(10, n_stocks) # Create difference between stocks
+N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)
+
+NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
+               sigma=rep(sigma,n_stocks),
+               cor=rep(re_cor,n_stocks),
+               recruit_model = 3,
+               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
+               sigma_vals = sigma_vals,
+               N1_pars = N1_pars)#,
+               #NAA_where = basic_info$NAA_where)
+
+# recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
+# recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
+# recruit_model = 3: Beverton-Holt stock-recruitment with annual recruitments as random effects
+# recruit_model = 4: Ricker stock-recruitment with annual recruitments as random effects
+
+# 1. recruit_pars: a list (length = n_stocks) of vectors of initial parameters for recruitment model. 
+# If $recruit_model is 3 (B-H) or 4 (Ricker), parameters are "alpha" and "beta".
+
+# 2. sigma_vals: Initial standard deviation values to use for the NAA deviations. Values are not used if recruit_model = 1 
+# If sigma="rec": must be a list (length = n_stocks) of single values
+# If sigma="rec+1": a list (length = n_stocks) of 2 values must be specified. First is for the first age class (recruits), second is for all other ages.
+
+input <- prepare_wham_input(basic_info = basic_info, 
+                            selectivity = sel, 
+                            M = M, 
+                            NAA_re = NAA_re, 
+                            move = move,
+                            catch_info = catch_info, 
+                            index_info = index_info, 
+                            F = F_info)
+
+# IMPORTANT!#
+# This appears to be due to the initial F value for the Newton iterations to find F40. The default value is too high for long-lived fish
+input$data$FXSPR_init[] <- 0.01 # change to a low value
+
+random = input$random # check what processes are random effects
+input$random = NULL # so inner optimization won't change simulated RE
+om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
+# Note: do.fit must be FALSE (no modeling fitting yet)
+
+om_with_data <- update_om_fn(om, seed = 123, random = random)
+
+assess.interval <- 3 # 
+base.years      <- year_start:year_end # Burn-in period
+first.year      <- head(base.years,1)
+terminal.year   <- tail(base.years,1)
+assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)
+
+mods <- list() # Create a list to save MSE outputs
+
+NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
+# 
+# M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M
+
+mods[[1]] = loop_through_fn(om = om_with_data,
+                            em_info = info, 
+                            random = random,
+                            M_em = M, # use OM M
+                            sel_em = sel, # use OM sel
+                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
+                            move_em = NULL,
+                            age_comp_em = "multinomial",
+                            # Here is the correct code: separate.em = FALSE also works for one-area model
+                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
+                                          do.move = FALSE, est.move = FALSE),
+                            assess_years = assess.years, 
+                            assess_interval = assess.interval, 
+                            base_years = base.years,
+                            year.use = 20, # number of years of data you want to use in the assessment model
+                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
+                            seed = 123,
+                            save.sdrep = FALSE, 
+                            save.last.em = TRUE,
+                            FXSPR_init = 0.01) # IMPORTANT!
+
+```
 

### 2025-04-15 Scenarios.Rmd
push test from container
updates survey timing to 2024

@@
 
-cod assessment or redfish for these
+cod assessment or redfish for these (Brian supplied redfish!)
 
@@
 # from a local surdat pull 
-# fracyr <- survdat_nobio$survdat |> #file from last May, could update
+# survdat_nobio <- readRDS(here::here("localhugefiles/survdat_nolength.rds"))
+# fracyr <- survdat_nobio$survdat |> #file from April 2025
 #   dplyr::mutate(DOY = lubridate::yday(EST_TOWDATE)) |>
@@
 #   dplyr::group_by(SEASON) |>
-#   dplyr::summarise(medDOY = median(DOY), 
+#   dplyr::summarise(medDOY = median(DOY, na.rm=T),
 #                    fracyrmed = medDOY/365)
+# fracyr
+# # A tibble: 2 × 3
+#   SEASON medDOY fracyrmed
+#   <chr>   <dbl>     <dbl>
+# 1 FALL     296.     0.812
+# 2 SPRING   108.     0.295
 

### 2025-04-15 Scenarios.Rmd
update for long lived (redfish-like)

@@
 # Selectivity Configuration
-fleet_pars <- c(5,1)
-index_pars <- c(2,1)
+# for logistic pars are a50 and 1/slope
+# approximate from redfish model
+fleet_pars <- c(9,1)
+index_pars <- list(c(5,1), c(3,1))
 sel <- list(model=rep("logistic",n_fleets+n_indices),
-            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
+            initial_pars=c(rep(list(fleet_pars),n_fleets),
+                           #rep(list(index_pars),n_indices)))
+                           index_pars))
 
+# redfish M is 0.05 but I'll leave this at 0.1 
 # M Configuration
@@
 
-# Set para. for B-H function
-alpha <- 12
-beta  <- 1.5e-4
+# # Set para. for B-H function, not using
+# alpha <- 12
+# beta  <- 1.5e-4
 
@@
                cor=rep(re_cor,n_stocks),
-               recruit_model = 3,
-               recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
+               recruit_model = 2,
+               #recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
                sigma_vals = sigma_vals,

### 2025-04-15 Scenarios.Rmd
trying to add data poor scenario, breaking

@@
 
+Looks like ~0.2 is reasonable for redfish in ICELAND, higher here.
+
+Say its decent ~0.3
+
 Eff = 100
@@
 
-Project at F40%
+Project at F40% (75% F40% is HCR 1)
 
@@
 cod assessment or redfish for these (Brian supplied redfish!)
+
+Keep multinomial
+
+Single fleet logistic selectivity: shift one age to the right from maturity ogive
+
+Different selectivity by spring and fall survey: mimic some long lived species
 
@@
                             catch_info = list(catch_cv = 0.1, catch_Neff = 100),
-                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = c(0.295, 0.812), q = 0.2),
+                            index_info = list(index_cv = 0.3, index_Neff = 100, fracyr_indices = c(0.295, 0.812), q = 0.2),
                             fracyr_spawn = 0.625)
@@
 
+# -------------------
+# IMPORTANT!#
+# -------------------
+# Change initial numbers at age in the input for OM
+# ini.NAA <- matrix(NA, n_ages, n_stocks)
+# ini.NAA[,1] <- c(20:1) # Use actually initial numbers at age from assessment model?
+# 
+# input$par$log_N1[] = 0
+# for (i in 1:n_regions) {
+#   input$par$log_N1[i,i,] = log(ini.NAA[,i])
+# }
+
 random = input$random # check what processes are random effects
@@
                             base_years = base.years,
-                            year.use = 20, # number of years of data you want to use in the assessment model
+                            year.use = 30, # number of years of data you want to use in the assessment model
                             add.years = TRUE, # extends assessment time series instead of moving window of year.use years
@@
 
-Keep multinomial
+```{r}
+agg_index_sigma = input$data$agg_index_sigma
+agg_index_sigma[31:60,] = 0.6 # Increase CV for both survey indices in the feedback period
+index_Neff = input$data$index_Neff
+index_Neff[31:60,] = 50 # Decrease ESS for both survey indices in the feedback period
 
-Single fleet logistic selectivity: shift one age to the right from maturity ogive
+#alternate years fall and spring surveys
+remove_agg = TRUE # remove a aggregate index for some years
+remove_agg_pointer = c(1,2) # both
+remove_agg_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
+remove_paa = TRUE # Also remove age comp for that index 
+remove_paa_pointer = c(1,2) # both
+remove_paa_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
 
-Different selectivity by spring and fall survey: mimic some long lived species
+input <- update_input_index_info(input, agg_index_sigma, index_Neff,
+                                 remove_agg, remove_agg_pointer, remove_agg_years,
+                                 remove_paa, remove_paa_pointer, remove_paa_years) # Update input file
+
+agg_catch_sigma = input$data$agg_catch_sigma
+agg_catch_sigma[31:60,] = 0.2 #double catch CV in the feedback period
+catch_Neff = input$data$catch_Neff
+catch_Neff[31:60] = 50
+
+input <- update_input_catch_info(input, agg_catch_sigma, catch_Neff)
+
+```
+
+Generate new om and dataset, keep same assessment interval as above
+
+```{r}
+om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
+om_with_data <- update_om_fn(om, seed = 123, random = random)
+```
+
+Test new EM
+
+```{r}
+
+mods[[2]] = loop_through_fn(om = om_with_data,
+                            em_info = info, 
+                            random = random,
+                            M_em = M, # use OM M
+                            sel_em = sel, # use OM sel
+                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
+                            move_em = NULL,
+                            age_comp_em = "multinomial",
+                            # Here is the correct code: separate.em = FALSE also works for one-area model
+                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
+                                          do.move = FALSE, est.move = FALSE),
+                            assess_years = assess.years, 
+                            assess_interval = assess.interval, 
+                            base_years = base.years,
+                            year.use = 30, # number of years of data you want to use in the assessment model
+                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
+                            seed = 123,
+                            save.sdrep = FALSE, 
+                            save.last.em = TRUE,
+                            FXSPR_init = 0.01) # IMPORTANT!
+
+```
+
+Visualize 2 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)
+
+This compares the base case,good surveys in the future (1) with degraded data, alternating year seasonal surveys with lower quality sampling (catch too) in the future (2)
+
+```{r}
+
+par(mfrow = c(1,2))
+SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
+plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB") 
+colors <- c("red","green","purple","orange")
+for (i in length(SSB_s1)) {
+  lines(SSB_s1[[i]], col = colors[i-1],lty = i)
+}
+legend("topleft",legend = paste0("EM ", 1:2), col = c("blue",colors),lty=1:5, cex=0.8)
+
+Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
+plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,30000), main = "OM Catch")
+colors <- c("red","green","purple","orange")
+for (i in length(Catch_s1)) {
+  lines(Catch_s1[[i]], col = colors[i-1],lty = i)
+}
+
+```
 
@@
 
+Do this later...
+
+For now, do
+
 100 replicates for each, try the parallel processing

### 2025-04-15 Scenarios.Rmd
update recsig, mean_rec_par in base

@@
 # Set sigma for NAA
+# add higher rec_sig similar to redfish WHAM
+rec_sig <- 1.0
 NAA_sig <- 0.2
 sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages"
+sigma_vals[,,1] = Rec_sig
 
@@
 
+# Set mean recruitment para. for each stock
+mean_rec_par <- list()
+for (i in 1:n_stocks) mean_rec_par[[i]] <- exp(log_N1[i])
+
 NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
@@
                recruit_model = 2,
-               #recruit_pars = rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
+               recruit_pars = mean_rec_par, # rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
                sigma_vals = sigma_vals,

### 2025-04-15 Scenarios.Rmd
Rec_sig that is

@@
 # add higher rec_sig similar to redfish WHAM
-rec_sig <- 1.0
+Rec_sig <- 1.0
 NAA_sig <- 0.2

### 2025-04-16 Scenarios.Rmd
Cheng's corrections

@@
 
+# --------------------------------------------------- #
+# ------- Do we need to change fracyr_spawn? -------- #
+#    Fraction of the year when spawning is now 0.625  #
+# --------------------------------------------------- #
+
 basic_info = info$basic_info # collect basic information
@@
 log_N1  <- rep(10, n_stocks) # Create difference between stocks
-N1_pars <- generate_ini_N1(log_N1,basic_info,ini.opt)
+
+# ------------------------------------------------------ #
+# ------------- Below code has been changed ------------ #
+# ------------------------------------------------------ #
+
+N1_pars <- generate_ini_N1(basic_info = basic_info, ini.opt = "equilibrium", log_N1 = log_N1, log_N1_F = 0.1)
+# WE MAY NEED TO HAVE A MORE SENSIBLE INITIAL NUMBERS AT AGE (now is using default: Number of age 1 = exp(10))
+
+# ------------------------------------------------------ #
+# ------------- Above code has been changed ------------ #
+# ------------------------------------------------------ #
 
@@
 mean_rec_par <- list()
-for (i in 1:n_stocks) mean_rec_par[[i]] <- exp(log_N1[i])
+for (i in 1:n_stocks) mean_rec_par[[i]] <- exp(log_N1[i]) #
+
+# ------------------------------------------------------ #
+# ------ WE MAY NEED A SENSIBLE MEAN RECREUITMENT ------ #
+# ------------------------------------------------------ #
 
@@
                                           do.move = FALSE, est.move = FALSE),
+                            # ------------------------------------------------------ #
+                            # - Below is needed when making changes on data quality- #
+                            # ------------------------------------------------------ #
+                            update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
+                            update_catch_info  = list(agg_catch_sigma = agg_catch_sigma, catch_Neff = catch_Neff), # Must have this!
+                            # ------------------------------------------------------ #
+                            # - Above is needed when making changes on data quality- #
+                            # ------------------------------------------------------ #
                             assess_years = assess.years, 
@@
 SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
-plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB") 
+plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB", lwd = 2) 
 colors <- c("red","green","purple","orange")
-for (i in length(SSB_s1)) {
-  lines(SSB_s1[[i]], col = colors[i-1],lty = i)
-}
-legend("topleft",legend = paste0("EM ", 1:2), col = c("blue",colors),lty=1:5, cex=0.8)
+lines(SSB_s1[[2]], col = colors[2],lty = 2, lwd = 2)
+legend("topleft",legend = paste0("EM ", 1:2), col = c("red","green"),lty=1:2, cex=0.8)
 
 Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
-plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,30000), main = "OM Catch")
+plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,15000), main = "OM Catch", lwd = 2)
 colors <- c("red","green","purple","orange")
-for (i in length(Catch_s1)) {
-  lines(Catch_s1[[i]], col = colors[i-1],lty = i)
-}
+lines(Catch_s1[[2]], col = colors[2],lty = 2, lwd = 2)
 

### 2025-04-16 docs/index.Rmd
trying long lived MSE setup, currently breaking
adds output of base case and data poor scenarios

@@
 
-<a href="https://noaa-edab.github.io/WHAMmse_longlived/WorkflowDecisions.html" target="_blank"> Documenting long lived MSE workflow - April 2025 </a>
+<a href="https://noaa-edab.github.io/WHAMmse_longlived/MSEsetup.html" target="_blank"> Documenting long lived MSE workflow - April 2025 </a>
 
-<a href="https://noaa-edab.github.io/WHAMmse_longlived/Results.html" target="_blank"> Initial results - April 2025 </a>
+<a href="https://noaa-edab.github.io/WHAMmse_longlived/Scenarios.html" target="_blank"> Single run of base and data degradation scenarios - April 2025 </a>

### 2025-04-16 ParallelTest.Rmd
start trying parallel processing

--- 2025-04-16 Scenarios.Rmd
@@
 ---
-title: "MSE Scenarios for short and long lived"
+title: "Parallel Test"
 author: "Sarah Gaichas"
@@
 
-## Setup for MSEs
+## Test parallel script with [Scenarios](https://noaa-edab.github.io/WHAMmse_longlived/Scenarios.html)
 
-Single area
+See Cheng's example https://lichengxue.github.io/whamMSE/05.Parallel-Computing.html
 
-Two surveys spring and fall
+First set up the base model, all the way up to the blank mods list, but not sampling data yet
 
-Testing alternating spring and fall
-
-Baseline is all data all the time,
-
-CV = empirical value from stockeff across short and long lived species
-
-Looks like ~0.2 is reasonable for redfish in ICELAND, higher here.
-
-Say its decent ~0.3
-
-Eff = 100
-
-Catch CV = 10%
-
-Eff = 100
-
-Project at F40% (75% F40% is HCR 1)
-
-Assessment frequency 3 years
-
-IID NAA
-
-30 year spin up 30 year loop (10 assessment cycles)
-
-F scenario for spin up? high followed by lower? Yes
-
-Double Fmsy for 15 years and then Fmsy for 15 years
-
-F.year1 = 0.1, Fhist = "F-H-L", Fmax = 2.5, Fmin = 1, changepoint = 0.5
-
-sigma squared for NAA and recruitment?
-
-cod assessment or redfish for these (Brian supplied redfish!)
-
-Keep multinomial
-
-Single fleet logistic selectivity: shift one age to the right from maturity ogive
-
-Different selectivity by spring and fall survey: mimic some long lived species
-
-I think this script does this
-
-```{r}
-
+```{r base setup 3yr}
 library(wham)
@@
 
-om_with_data <- update_om_fn(om, seed = 123, random = random)
+#om_with_data <- update_om_fn(om, seed = 123, random = random)
 
@@
 
+```
+
+Set up folders for results
+
+```{r}
+sub.dir = "Results"
+dir.create(file.path(getwd(), sub.dir), recursive = TRUE)
+
+library(doParallel)
+library(foreach)
+
+detectCores() # check how many cores available
+```
+
+Set up clusters to run multiple sims for each scenario
+
+```{r}
+cluster <- makeCluster(5) 
+registerDoParallel(cluster)
+
+foreach (i = 1:5) %dopar% {
+
+  library(wham)
+  library(whamMSE)
+
+  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
+
   NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
-# 
-# M_em <- list(model="constant",initial_means=array(0.1, dim = c(n_stocks,n_regions,n_ages))) # this is misspecified M
 
-mods[[1]] = loop_through_fn(om = om_with_data,
+  mod = loop_through_fn(om = om_with_data,
                             em_info = info, 
@@
                             add.years = TRUE, # extends assessment time series instead of moving window of year.use years
-                            seed = 123,
+                            seed = 123+i,
                             save.sdrep = FALSE, 
@@
 
+  saveRDS(mod, file.path(sub.dir,sprintf("Mod1_%03d.RDS",i)))
+
+}
+
+stopCluster(cluster)
+
 ```
-
-## Degraded data quality scenarios:
-
-Alternate spring and fall, double CV and half Eff (50)
-
-Fishery data, double CV (0.2) and half Eff (50)
+Now try the same thing (base case) changing the assessment frequency
 
 ```{r}
-agg_index_sigma = input$data$agg_index_sigma
-agg_index_sigma[31:60,] = 0.6 # Increase CV for both survey indices in the feedback period
-index_Neff = input$data$index_Neff
-index_Neff[31:60,] = 50 # Decrease ESS for both survey indices in the feedback period
+assess.interval <- 6 #
 
-#alternate years fall and spring surveys
-remove_agg = TRUE # remove a aggregate index for some years
-remove_agg_pointer = c(1,2) # both
-remove_agg_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
-remove_paa = TRUE # Also remove age comp for that index 
-remove_paa_pointer = c(1,2) # both
-remove_paa_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
+cluster <- makeCluster(5) 
+registerDoParallel(cluster)
 
-input <- update_input_index_info(input, agg_index_sigma, index_Neff,
-                                 remove_agg, remove_agg_pointer, remove_agg_years,
-                                 remove_paa, remove_paa_pointer, remove_paa_years) # Update input file
+foreach (i = 1:5) %dopar% {
 
-agg_catch_sigma = input$data$agg_catch_sigma
-agg_catch_sigma[31:60,] = 0.2 #double catch CV in the feedback period
-catch_Neff = input$data$catch_Neff
-catch_Neff[31:60] = 50
+  library(wham)
+  library(whamMSE)
 
-input <- update_input_catch_info(input, agg_catch_sigma, catch_Neff)
+  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
 
-```
+  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
 
-Generate new om and dataset, keep same assessment interval as above
-
-```{r}
-om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
-om_with_data <- update_om_fn(om, seed = 123, random = random)
-```
-
-Test new EM
-
-```{r}
-
-mods[[2]] = loop_through_fn(om = om_with_data,
+  mod = loop_through_fn(om = om_with_data,
                             em_info = info, 
@@
                                           do.move = FALSE, est.move = FALSE),
-                            # ------------------------------------------------------ #
-                            # - Below is needed when making changes on data quality- #
-                            # ------------------------------------------------------ #
-                            update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
-                            update_catch_info  = list(agg_catch_sigma = agg_catch_sigma, catch_Neff = catch_Neff), # Must have this!
-                            # ------------------------------------------------------ #
-                            # - Above is needed when making changes on data quality- #
-                            # ------------------------------------------------------ #
                             assess_years = assess.years, 
@@
 
+  saveRDS(mod, file.path(sub.dir,sprintf("Mod2_%03d.RDS",i)))
+
+}
+
+stopCluster(cluster)
 ```
-
-Visualize 2 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)
-
-This compares the base case,good surveys in the future (1) with degraded data, alternating year seasonal surveys with lower quality sampling (catch too) in the future (2)
-
-```{r}
-
-par(mfrow = c(1,2))
-SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
-plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB", lwd = 2) 
-colors <- c("red","green","purple","orange")
-lines(SSB_s1[[2]], col = colors[2],lty = 2, lwd = 2)
-legend("topleft",legend = paste0("EM ", 1:2), col = c("red","green"),lty=1:2, cex=0.8)
-
-Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
-plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,15000), main = "OM Catch", lwd = 2)
-colors <- c("red","green","purple","orange")
-lines(Catch_s1[[2]], col = colors[2],lty = 2, lwd = 2)
-
-```
-
-## Empirical management procedure
-
-Index based method:  Ismooth averaging most recent 3 years of whatever surveys (rescaled to mean of each using Ismooth package)
-
-Ismooth package is here https://github.com/cmlegault/PlanBsmooth
-
-Do this later...
-
-For now, do
-
-100 replicates for each, try the parallel processing
-
-* Base scenario 
-* Alternating surveys and less age comps with 3 year assessments
-* Empirical management procedures (later)

### 2025-04-16 ParallelTest.Rmd
Cheng's fix and sketch in data poor scenarios

@@
 
-```{r}
+```{r, eval=FALSE}
 sub.dir = "Results"
@@
 
+base.years      <- year_start:year_end # Burn-in period
+first.year      <- head(base.years,1)
+terminal.year   <- tail(base.years,1)
+assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)
+
 cluster <- makeCluster(5) 
@@
 ```
+
+That isn't working
+
+Try the low data example that worked individually, first the setup
+
+```{r}
+agg_index_sigma = input$data$agg_index_sigma
+agg_index_sigma[31:60,] = 0.6 # Increase CV for both survey indices in the feedback period
+index_Neff = input$data$index_Neff
+index_Neff[31:60,] = 50 # Decrease ESS for both survey indices in the feedback period
+
+#alternate years fall and spring surveys
+remove_agg = TRUE # remove a aggregate index for some years
+remove_agg_pointer = c(1,2) # both
+remove_agg_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
+remove_paa = TRUE # Also remove age comp for that index 
+remove_paa_pointer = c(1,2) # both
+remove_paa_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
+
+input <- update_input_index_info(input, agg_index_sigma, index_Neff,
+                                 remove_agg, remove_agg_pointer, remove_agg_years,
+                                 remove_paa, remove_paa_pointer, remove_paa_years) # Update input file
+
+agg_catch_sigma = input$data$agg_catch_sigma
+agg_catch_sigma[31:60,] = 0.2 #double catch CV in the feedback period
+catch_Neff = input$data$catch_Neff
+catch_Neff[31:60] = 50
+
+input <- update_input_catch_info(input, agg_catch_sigma, catch_Neff)
+
+om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
+
+```
+
+Now run in parallel, 3 year assessment
+
+```{r}
+assess.interval <- 3 #
+
+base.years      <- year_start:year_end # Burn-in period
+first.year      <- head(base.years,1)
+terminal.year   <- tail(base.years,1)
+assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)
+
+cluster <- makeCluster(5) 
+registerDoParallel(cluster)
+
+foreach (i = 1:5) %dopar% {
+
+  library(wham)
+  library(whamMSE)
+
+  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
+
+  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
+
+  mod = loop_through_fn(om = om_with_data,
+                            em_info = info, 
+                            random = random,
+                            M_em = M, # use OM M
+                            sel_em = sel, # use OM sel
+                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
+                            move_em = NULL,
+                            age_comp_em = "multinomial",
+                            # Here is the correct code: separate.em = FALSE also works for one-area model
+                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
+                                          do.move = FALSE, est.move = FALSE),
+                            assess_years = assess.years, 
+                            assess_interval = assess.interval, 
+                            base_years = base.years,
+                            year.use = 30, # number of years of data you want to use in the assessment model
+                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
+                            seed = 123+i,
+                            save.sdrep = FALSE, 
+                            save.last.em = TRUE,
+                            FXSPR_init = 0.01) # IMPORTANT!
+
+  saveRDS(mod, file.path(sub.dir,sprintf("Mod3_%03d.RDS",i)))
+
+}
+
+stopCluster(cluster)
+
+```
+
+And degraded data with a 6 year assessment
+
+```{r}
+assess.interval <- 6 #
+
+base.years      <- year_start:year_end # Burn-in period
+first.year      <- head(base.years,1)
+terminal.year   <- tail(base.years,1)
+assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)
+
+cluster <- makeCluster(5) 
+registerDoParallel(cluster)
+
+foreach (i = 1:5) %dopar% {
+
+  library(wham)
+  library(whamMSE)
+
+  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
+
+  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
+
+  mod = loop_through_fn(om = om_with_data,
+                            em_info = info, 
+                            random = random,
+                            M_em = M, # use OM M
+                            sel_em = sel, # use OM sel
+                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
+                            move_em = NULL,
+                            age_comp_em = "multinomial",
+                            # Here is the correct code: separate.em = FALSE also works for one-area model
+                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
+                                          do.move = FALSE, est.move = FALSE),
+                            assess_years = assess.years, 
+                            assess_interval = assess.interval, 
+                            base_years = base.years,
+                            year.use = 30, # number of years of data you want to use in the assessment model
+                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
+                            seed = 123+i,
+                            save.sdrep = FALSE, 
+                            save.last.em = TRUE,
+                            FXSPR_init = 0.01) # IMPORTANT!
+
+  saveRDS(mod, file.path(sub.dir,sprintf("Mod4_%03d.RDS",i)))
+
+}
+
+stopCluster(cluster)
+
+```

### 2025-04-16 RunParallel4scen.R
update parallel test and draft script for 4x100 sims

@@ -1,21 +1,8 @@
----
-title: "Parallel Test"
-author: "Sarah Gaichas"
-date: "`r Sys.Date()`"
-output: html_document
----
+# script to run in parallel on container
+# based on tests in ParallelTest.Rmd
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+# Set up base case with 3 year assessment interval
 
-## Test parallel script with [Scenarios](https://noaa-edab.github.io/WHAMmse_longlived/Scenarios.html)
-
-See Cheng's example https://lichengxue.github.io/whamMSE/05.Parallel-Computing.html
-
-First set up the base model, all the way up to the blank mods list, but not sampling data yet
-
-```{r base setup 3yr}
 library(wham)
 library(whamMSE)
 
@@ -135,7 +122,7 @@ NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
                recruit_pars = mean_rec_par, # rep(list(c(alpha,beta)),n_stocks), # assume same B-H s-r functions for all stocks
                sigma_vals = sigma_vals,
                N1_pars = N1_pars)#,
-               #NAA_where = basic_info$NAA_where)
+#NAA_where = basic_info$NAA_where)
 
 # recruit_model = 1: estimating annual recruitments as fixed effects or a random walk if NAA_re$sigma specified
 # recruit_model = 2: estimating a mean recruitment with annual recruitments as random effects
@@ -190,28 +177,22 @@ assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = asse
 
 mods <- list() # Create a list to save MSE outputs
 
-```
+## Make directory
 
-Set up folders for results
-
-```{r, eval=FALSE}
 sub.dir = "Results"
 dir.create(file.path(getwd(), sub.dir), recursive = TRUE)
 
 library(doParallel)
 library(foreach)
 
-detectCores() # check how many cores available
-```
+n_cores <- detectCores() # check how many cores available
 
+# Run base case, 3 year assessment interval, 100 reps
 
-Set up clusters to run multiple sims for each scenario
-
-```{r}
-cluster <- makeCluster(5) 
+cluster <- makeCluster(n_cores-1) 
 registerDoParallel(cluster)
 
-foreach (i = 1:5) %dopar% {
+foreach (i = 1:100) %dopar% {
   
   library(wham)
   library(whamMSE)
@@ -219,28 +200,28 @@ foreach (i = 1:5) %dopar% {
[…]

### 2025-04-16 ParallelTest.Rmd
update info for degraded models

@@
 ```
+
 Now try the same thing (base case) changing the assessment frequency
@@
 ```
-
-That isn't working
 
@@
                             add.years = TRUE, # extends assessment time series instead of moving window of year.use years
-                            seed = 123+i,
+                            seed = 123,
                             save.sdrep = FALSE, 

### 2025-04-16 RunParallel4scen.R

@@ -339,12 +339,20 @@ foreach (i = 1:100) %dopar% {
                         # Here is the correct code: separate.em = FALSE also works for one-area model
                         em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                       do.move = FALSE, est.move = FALSE),
+                        # ------------------------------------------------------ #
+                        # - Below is needed when making changes on data quality- #
+                        # ------------------------------------------------------ #
+                        update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
+                        update_catch_info  = list(agg_catch_sigma = agg_catch_sigma, catch_Neff = catch_Neff), # Must have this!
+                        # ------------------------------------------------------ #
+                        # - Above is needed when making changes on data quality- #
+                        # ------------------------------------------------------ #
                         assess_years = assess.years, 
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 30, # number of years of data you want to use in the assessment model
                         add.years = TRUE, # extends assessment time series instead of moving window of year.use years
-                        seed = 123+i,
+                        seed = 123,
                         save.sdrep = FALSE, 
                         save.last.em = TRUE,
                         FXSPR_init = 0.01) # IMPORTANT!
@@ -390,12 +398,20 @@ foreach (i = 1:100) %dopar% {
                         # Here is the correct code: separate.em = FALSE also works for one-area model
                         em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                       do.move = FALSE, est.move = FALSE),
+                        # ------------------------------------------------------ #
+                        # - Below is needed when making changes on data quality- #
+                        # ------------------------------------------------------ #
+                        update_index_info  = list(agg_index_sigma = agg_index_sigma, index_Neff = index_Neff), # Must have this!
+                        update_catch_info  = list(agg_catch_sigma = agg_catch_sigma, catch_Neff = catch_Neff), # Must have this!
+                        # ------------------------------------------------------ #
+                        # - Above is needed when making changes on data quality- #
+                        # ------------------------------------------------------ #
                         assess_years = assess.years, 
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 30, # number of years of data you want to use in the assessment model
                         add.years = TRUE, # extends assessment time series instead of moving window of year.use years
-                        seed = 123+i,
+                        seed = 123,
[…]

### 2025-04-16 RunParallel4scen.R
remove unnecessary stuff

@@ -6,24 +6,6 @@
 library(wham)
 library(whamMSE)
 
-main.dir = here::here()
-
-# get median DOY for each seasonal survey
-# from a local surdat pull 
-# survdat_nobio <- readRDS(here::here("localhugefiles/survdat_nolength.rds"))
-# fracyr <- survdat_nobio$survdat |> #file from April 2025
-#   dplyr::mutate(DOY = lubridate::yday(EST_TOWDATE)) |>
-#   dplyr::select(SEASON, DOY) |> dplyr::distinct() |>
-#   dplyr::group_by(SEASON) |>
-#   dplyr::summarise(medDOY = median(DOY, na.rm=T),
-#                    fracyrmed = medDOY/365)
-# fracyr
-# # A tibble: 2 × 3
-#   SEASON medDOY fracyrmed
-#   <chr>   <dbl>     <dbl>
-# 1 FALL     296.     0.812
-# 2 SPRING   108.     0.295
-
 year_start  <- 1  # starting year in the burn-in period
 year_end    <- 30  # end year in the burn-in period
 MSE_years   <- 30     # number of years in the feedback loop

### 2025-04-25 RunParallel4scen.R
increment seed and rerun 3 and 4

@@ -334,7 +334,7 @@ foreach (i = 1:100) %dopar% {
                         base_years = base.years,
                         year.use = 30, # number of years of data you want to use in the assessment model
                         add.years = TRUE, # extends assessment time series instead of moving window of year.use years
-                        seed = 123,
+                        seed = 123+i,
                         save.sdrep = FALSE, 
                         save.last.em = TRUE,
                         FXSPR_init = 0.01) # IMPORTANT!
@@ -393,7 +393,7 @@ foreach (i = 1:100) %dopar% {
                         base_years = base.years,
                         year.use = 30, # number of years of data you want to use in the assessment model
                         add.years = TRUE, # extends assessment time series instead of moving window of year.use years
-                        seed = 123,
+                        seed = 123+i,
                         save.sdrep = FALSE, 
                         save.last.em = TRUE,
                         FXSPR_init = 0.01) # IMPORTANT!

### 2025-04-28 ParallelTest.Rmd
fix bug in this demo too

@@
 ```{r, eval=FALSE}
-sub.dir = "Results"
+sub.dir = "Results_test"
 dir.create(file.path(getwd(), sub.dir), recursive = TRUE)

### 2025-04-28 PlotScenarios.Rmd
test plots

---
title: "Plot Scenarios"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Test plots with small subset

The first 5 runs of baseline with 3 year assessment are in Results_test

What should we plot? SSB, Catch, F, Rec? How often the stock is overfished, interannual variation in catch?

Read them in and build an object to plot

```{r}
# use the purrr library

# function to pull the pieces from a single rds file

d.name <- here::here("Results_test")
modnames <- list.files(d.name)

modpaths <- paste0(d.name, "/", modnames)

getmodts <- function(path){

  mod <- readRDS(here::here(path))
  modpath <- stringr::str_split(path, "/", simplify = TRUE)
  filename <- modpath[length(modpath)]
  modname <- stringr::str_split(filename, ".RDS", simplify = TRUE)[1]
  sceniter <- stringr::str_split(modname, "_", simplify = TRUE)
  scen <- sceniter[1]
  iter <- sceniter[2]

  omSSB <- mod$om$rep$SSB
  omcatch <- mod$om$rep$pred_catch

   # return model indices as a dataframe
  out <- data.frame(modname = modname,
                    scen = scen,
                    iter = iter,
                    year = 1:length(omSSB),
                    omSSB = omSSB,
                    omcatch = omcatch
                    )

  return(out)
}

# apply the function over all rds in the directory

testdat <- purrr::map_dfr(modpaths, getmodts)

# plot SSB and catch for each model iteration
# needs patchwork library

library(patchwork)

SSB <- ggplot2::ggplot(testdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omSSB, colour = iter)) +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen) +
  ggplot2::theme(legend.position = "none")

Catch <- ggplot2::ggplot(testdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omcatch, colour = iter))+
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen) +
  ggplot2::theme(legend.position = "none")

SSB + Catch

```

## A test with all 400

Its a start

```{r}
d.name <- here::here("Results")
modnames <- list.files(d.name)

modpaths <- paste0(d.name, "/", modnames)

# apply the function over all rds in the directory

realdat <- purrr::map_dfr(modpaths, getmodts)

# plot SSB and catch for each model iteration
# needs patchwork library

library(patchwork)

scennames <- c(`Mod1` = "Base 3 yr",
               `Mod2` = "Base 6 yr",
               `Mod3` = "Degraded data 3 yr",
               `Mod4` = "Degraded data 6 yr")

SSB <- ggplot2::ggplot(realdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omSSB, colour = iter),  alpha = 0.3) + #
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen, labeller = ggplot2::as_labeller(scennames)) +
  ggplot2::theme(legend.position = "none")

Catch <- ggplot2::ggplot(realdat) +
  ggplot2::geom_line(ggplot2::aes(x=year, y=omcatch, colour = iter), alpha = 0.3) + #
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(~scen, labeller = ggplot2::as_labeller(scennames)) +
  ggplot2::theme(legend.position = "none")

SSB

Catch
```

### 2025-04-30 PlotScenarios.Rmd
added Cheng's code from 4/29 but not 4/30, email cut off

@@
 ```
+Add in Cheng's new plotting functions
+
+```{r}
+source(here::here("helper.R"))
+```
+
+Cheng's example modified from `MSE_plot_code.R`
+
+```{r}
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(fmsb)
+library(ggpubr)
+library(gridExtra)
+library(rmarkdown)
+
+main.dir = getwd()
+
+sub.dir <- 'Report'
+output_dir <- file.path(main.dir,sub.dir)
+
+if(!dir.exists(output_dir)) {
+  dir.create(output_dir, recursive = TRUE)
+}
+
+model_nums <- c(1,4)
+nsim <- 50 # number of simulations/seed
+
+mods <- vector("list", nsim)
+
+for (r in 1:nsim) {
+  mods[[r]] <- vector("list", length(model_nums))
+  names(mods[[r]]) <- paste0("Mod", model_nums)
+
+  for (m in model_nums) {
+    file_path <- file.path("Results/", sprintf("Mod%d_%03d.RDS", m, r))
+    mods[[r]][[paste0("Mod", m)]] <- readRDS(file_path)
+  }
+}
+
+if (!is.list(mods[[1]][[1]][[1]])) is.nsim = FALSE else is.nsim = TRUE
+
+cat("\nPloting SSB Time Series...\n")
+
+plot_ssb_time_series(mods, is.nsim, main.dir, sub.dir, var = "SSB",
+                     width = 10, height = 7, dpi = 300,
+                     new_model_names = NULL) # you can change the Model name here
+
+cat("\nPloting Fbar Time Series...\n")
+
+plot_fbar_time_series(mods, is.nsim, main.dir, sub.dir, var = "Fbar",
+                      width = 10, height = 7, dpi = 300,
+                      new_model_names = NULL)
+
+cat("\nPloting Catch Time Series...\n")
+
+plot_catch_time_series(mods, is.nsim, main.dir, sub.dir, var = "Catch",
+                       width = 10, height = 7, dpi = 300,
+                       new_model_names = NULL)
+
+cat("\nPloting SSB Performance Over Last n Years...\n")
+
+# Long-term performance #
+# Default = 5 years #
+
+plot_ssb_performance(mods, is.nsim, main.dir, sub.dir, var = "SSB",
+                     width = 10, height = 7, dpi = 300,
+                     new_model_names = NULL,
+                     use.n.years = NULL)
+
+cat("\nPloting Fbar Performance Over Last n Years...\n")
+
+plot_fbar_performance(mods, is.nsim, main.dir, sub.dir, var = "Fbar",
+                      width = 10, height = 7, dpi = 300,
+                      new_model_names = NULL,
+                      use.n.years = NULL)
+
+cat("\nPloting Catch Performance Over Last n Years...\n")
+
+plot_catch_performance(mods, is.nsim, main.dir, sub.dir, var = "Catch",
+                       width = 10, height = 7, dpi = 300,
+                       new_model_names = NULL,
+                       use.n.years = NULL)
+
+cat("\nPloting SSB Performance Over First n Years...\n")
+
+# -----------------------
+# Short-term performance
+# -----------------------
+# summarize first 10 years in the feedback period
+start.years = 31 
+use.n.years = 10
+
+plot_ssb_performance2(mods, is.nsim, main.dir, sub.dir, var = "SSB",
+                      width = 10, height = 7, dpi = 300,
+                      new_model_names = NULL,
+                      use.n.years = use.n.years,
+                      start.years = start.years)
+
+cat("\nPloting Fbar Performance Over First n Years...\n")
+
+plot_fbar_performance2(mods, is.nsim, main.dir, sub.dir, var = "Fbar",
+                       width = 10, height = 7, dpi = 300,
+                       new_model_names = NULL,
+                       use.n.years = use.n.years,
+                       start.years = start.years)
+
+cat("\nPloting Catch Performance Over First n Years...\n")
+
+plot_catch_performance2(mods, is.nsim, main.dir, sub.dir, var = "Catch",
+                        width = 10, height = 7, dpi = 300,
+                        new_model_names = NULL,
+                        use.n.years = use.n.years,
+                        start.years = start.years)
+
+cat("\nPloting SSB Status and Probability of Overfished Over Last n Years...\n")
+
+plot_ssb_status(mods, is.nsim, main.dir, sub.dir, var = "SSB_status",
+                width = 10, height = 7, dpi = 300,
+                new_model_names = NULL,
+                use.n.years = NULL)
+
+cat("\nPloting Fbar Status and Probability of Overfishing Over Last n Years...\n")
+
+plot_fbar_status(mods, is.nsim, main.dir, sub.dir, var = "Fbar_status",
+                 width = 10, height = 7, dpi = 300,
+                 new_model_names = NULL,
+                 use.n.years = NULL)
+
+cat("\nPloting SSB Status and Probability of Overfished Over First n Years...\n")
+
+# -----------------------
+# Short-term performance
+# -----------------------
+# summarize first 10 years in the feedback period
+start.years = 31 
+use.n.years = 10
+
+plot_ssb_status2(mods, is.nsim, main.dir, sub.dir, var = "SSB_status",
+                 width = 10, height = 7, dpi = 300,
+                 new_model_names = NULL,
+                 use.n.years = use.n.years,
+                 start.years = start.years)
+
+cat("\nPloting Fbar Status and Probability of Overfishing Over First n Years...\n")
+
+plot_fbar_status2(mods, is.nsim, main.dir, sub.dir, var = "Fbar_status",
+                  width = 10, height = 7, dpi = 300,
+                  new_model_names = NULL,
+                  use.n.years = use.n.years,
+                  start.years = start.years)
+
+```
